---
title: "yelp_reviews"
author: "Huey"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(stringi)
library(hash)

setwd("~/code/yelp")
```

# Introduction 
Are there differences in rating ethnic restaurants between people who are of
that ethnicity vs. not? For example, do Asians rate Asian restaurants differently
than non-Asians?

In this exploration, we look at Mexican restaurants and Chinese restaurants and
try to see if there are any differences. 

## Load businesses
The Yelp data contains 2.2M reviews and 591K tips by 552K users for 77K businesses.
The cities included are Edinburgh (UK), Karlsruhe (Germany), Montreal and Waterloo (Canada),
and Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison (US).
First, let's load the Yelp businesses. I used the JSON to CSV converter in https://github.com/Yelp/dataset-examples to make it easier for me to load.

```{r load data}
businesses = read_csv("data/yelp_academic_dataset_business.csv")
reviews = read_csv("data/yelp_academic_dataset_review.csv")
users = read_csv("data/yelp_academic_dataset_user.csv")
users$name = tolower(users$name)

latino.names = read.csv("popular_names_latino.csv", 
                        encoding = "latin1", stringsAsFactors = FALSE)
latino.names$Male.Name = stri_trans_general(latino.names$Male.Name, 
                                            'latin-ascii')
latino.names$Female.Name = stri_trans_general(latino.names$Female.Name, 
                                              'latin-ascii')

latino.names = as.data.frame(c(latino.names$Male.Name, latino.names$Female.Name), stringsAsFactors = FALSE)
names(latino.names) = "name"

latino.names.exclude = read.csv("popular_names_latino_exclude.csv", stringsAsFactors = FALSE)

latino.names = anti_join(latino.names, latino.names.exclude, by = "name")
latino.names$name = tolower(latino.names$name)
```

When classifying ethnicity, it's good to get a rough sense of how many users were
classified as Latino or not:

```{r latino users}
latino.users = semi_join(users, latino.names, by = "name")
non.latino.users = anti_join(users, latino.names, by = "name")

print(paste("Latino:", nrow(latino.users)))
print(paste("Non-Latino:" , nrow(non.latino.users)))
```

Less than 3% of users in the dataset have been classified as Latino.

Below is a function to get all the top n restaurants of a certain ethnicity and
bucket the user reviews into members of that ethnicity or not. We then run a 
t-test applying the Bonferroni correction to find which differences are 
statistically significant. 

```{r utility function}
GetRestaurantsWithDiffRatings = function(ethnicity, n.restaurants, eth.names) {
  eth.restaurants = businesses %>% filter(grepl(ethnicity, categories)) %>% filter(grepl('Restaurants', categories)) %>% arrange(desc(review_count))
  restaurants = eth.restaurants[1:n.restaurants, ]
  restaurant.reviews = inner_join(restaurants, reviews, by = "business_id")
  restaurant.reviews.users = inner_join(restaurant.reviews, users, by = "user_id")
  
  p.value.cutoff = 0.05 / n.restaurants

  restaurant_names = c()
  eth.ratings = c()
  non.eth.ratings = c()
  num.eth.reviews = c()
  num.non.eth.reviews = c()

  for (id in restaurants$business_id) {
    subset = filter(restaurant.reviews.users, business_id == id)
    eth.reviews = semi_join(subset, eth.names, c("name.y" = "name"))
    non.eth.reviews = anti_join(subset, eth.names, c("name.y" = "name"))
    
    if (length(eth.reviews$stars.y) <= 1 || length(non.eth.reviews$stars.y) <= 1) next  # Prevent errors to t-test
    
    try({
      alt = "two.sided"
      eth.mean = mean(eth.reviews$stars.y)
      non.eth.mean = mean(non.eth.reviews$stars.y)
      if (eth.mean < non.eth.mean) {
        alt = "less"
      }
      if (eth.mean > non.eth.mean) {
        alt = "greater"
      }
      res = t.test(eth.reviews$stars.y, non.eth.reviews$stars.y, alternative = alt)
      if (res$p.value < p.value.cutoff) {
        restaurant_names = c(restaurant_names, subset$name.x[1])
        eth.ratings = c(eth.ratings, eth.mean)
        non.eth.ratings = c(non.eth.ratings, non.eth.mean)
        num.eth.reviews = c(num.eth.reviews, nrow(eth.reviews))
        num.non.eth.reviews = c(num.non.eth.reviews, nrow(non.eth.reviews))
      }
    }
    )
  }
  return(data.frame(restaurant_names, eth.ratings, non.eth.ratings, num.eth.reviews, num.non.eth.reviews))
}
```

## Mexican Restaurants
The following table shows the restaurants where Latino ratings and non-Latino ratings
had statistically significant differences:

```{r latino data frame}
results = GetRestaurantsWithDiffRatings('Mexican', 100, latino.names)
results
```

For all of these, the sample size of Latino reviews is really small. It seems that
building a "racist Yelp" for Mexican restaurants would not be all that useful.

## Chinese Restaurants
To get Chinese-sounding names, we will compile a list of all pinyin sounds and
see if they exist in any first names. Unfortunately, this runs in quadratic time.

```{r chinese users}

load(file = "chinese_users.RData")

if (is.null(chinese.users)) {
  pinyin = read.csv("pinyin_sounds.csv", stringsAsFactors = FALSE)
  pinyin.pattern = paste(pinyin$sound, collapse ="|")
  
  exclude = "benji|ique|cque|arjun|axi|lexi|xine|jinx|quant|change|strong|cruise|chunky|suite|burgh fan|tx fan|eping|pping|lping"
  
  tictoc::tic()
  indexes = vector("logical", length = nrow(users))
  for (i in 1:nrow(users)) {
    if (length(grep(exclude, users$name[i])) > 0) next
    
    matches = grep(pinyin.pattern, users$name[i])
    indexes[i] = length(matches) > 0
  }
  tictoc::toc()
  chinese.users = users[indexes,]
  nrow(chinese.users)
  
  # save chinese.users in case it takes a long time
  save(chinese.users, file = "chinese_users.RData")
}

chinese.names = as.data.frame(unique(chinese.users$name), stringsAsFactors = FALSE)
names(chinese.names) = "name"
results = GetRestaurantsWithDiffRatings('Chinese', 100, chinese.names)
results
```



There are 7755 users classified as Chinese, or about 1.4% of the user population.
Per Wikipedia, Americans of Chinese descent consititute 1.2% of the total U.S.
popuplation as of 2010.

# Chinese Names Take 2

An alternative is to use the baby names from the following site:
http://www.top-100-baby-names-search.com/chinese-name-meanings.html
```{r chinese users take 2}
chinese.names2 = read.csv("top_100_baby_names_chinese.csv", stringsAsFactors = FALSE)
chinese.names2$name = tolower(chinese.names2$name)

chinese.names2 = chinese.names2[-124,]  # remove "juan"

chinese.users2 = semi_join(users, chinese.names2, by = "name")
nrow(chinese.users2)

results = GetRestaurantsWithDiffRatings('Chinese', 100, chinese.names2)
results
```

This yielded 2028 users and a different set of restaurants that had statistically
significantly different ratings. However, the number of reviews by Chinese people
are still to low to make any significant conclusions.

# East Asian Restaurants
What if we expand our search to Korean + Japanese + Chinese? Here we just use the 
baby names list for Chinese names.

```{r}
korean.names = read.csv("top_100_baby_names_korean.csv", stringsAsFactors = FALSE)
korean.names$name = tolower(korean.names$name)

japanese.names = read.csv("top_100_baby_names_japanese.csv", stringsAsFactors = FALSE)
japanese.names$name = tolower(japanese.names$name)

korean.users = semi_join(users, korean.names, by = "name")
japanese.users = semi_join(users, japanese.names, by = "name")

east.asian.names = as.data.frame(c(chinese.names2$name, japanese.names$name, korean.names$name), stringsAsFactors = FALSE)
names(east.asian.names) = c("name")
east.asian.users = semi_join(users, east.asian.names, by = "name")

results = GetRestaurantsWithDiffRatings('Chinese|Korean|Japanese', 100, east.asian.names)
results
```

# East Asian Users using pinyin for Chinese names
```{r east asian restaurants}
east.asian.names = as.data.frame(c(chinese.names$name, japanese.names$name, korean.names$name), stringsAsFactors = FALSE)
names(east.asian.names) = c("name")
east.asian.users = semi_join(users, east.asian.names)

results = GetRestaurantsWithDiffRatings('Chinese|Korean|Japanese', 100, east.asian.names)
results
```

Slightly higher rating (4.87 vs 4.67) for Gangnam Asian BBQ Dining from East Asians
than others in the data set.

# Conclusion
You can't construct a good "racist Yelp" from the data provided. If we had access
to last names, we might be able to include Asians or Latinos with Western first names,
which may change the results. 

## Restaurant preferences
What kind of restaurants do Latinos review?
What kind of restaurants do Asians review?

what's the distribution of review count on yelp?
```{r distribution of review count}
bw = function(x) diff(range(x)) / (2 * IQR(x) / length(x)^(1/3))
restaurants = businesses %>% filter(grepl('Restaurants', categories)) %>% arrange(desc(review_count))
ggplot(restaurants, aes(x = review_count)) + geom_histogram()
```

join latino users with restaurants, group by restaurants, count users
what are the names of the most reviewed restaurants
```{r review counts for ethnicity}
ReviewCountsForEthnicity = function(eth.names) {
  eth.users = semi_join(users, eth.names, by = "name")
  non.eth.users = anti_join(users, eth.names, by = "name")
  
  restaurant.reviews = inner_join(restaurants, reviews, by = "business_id")
  restaurant.reviews.eth.users = inner_join(restaurant.reviews, eth.users, by = "user_id")
  restaurant.reviews.non.eth.users = inner_join(restaurant.reviews, non.eth.users, by = "user_id")
  
  by.restaurant = group_by(restaurant.reviews.eth.users, business_id)
  eth.summary = summarise(by.restaurant, num.reviews = n()) %>% arrange(desc(num.reviews))
  eth.summary = inner_join(eth.summary, restaurants, by = "business_id") %>% select(business_id, name, num.reviews, categories)
  
  by.restaurant = group_by(restaurant.reviews.non.eth.users, business_id)
  non.eth.summary = summarise(by.restaurant, num.reviews = n()) %>% arrange(desc(num.reviews))
  non.eth.summary = inner_join(non.eth.summary, restaurants, by = "business_id") %>% select(business_id, name, num.reviews, categories)
  
  return(list(eth.summary, non.eth.summary))
}
```

```{r compile stats}
CompileStats = function(names, n) {
  l = list()
  summaries = ReviewCountsForEthnicity(names)
  n.eth.summaries = head(summaries[[1]], n)
  l[[1]] = n.eth.summaries
  
  n.non.eth.summaries = head(summaries[[2]], n)
  l[[2]] = n.non.eth.summaries
  
  l[[3]] = semi_join(n.eth.summaries, n.non.eth.summaries.id.name, by = "business_id")
  l[[4]] = anti_join(n.eth.summaries, n.non.eth.summaries, by = "business_id")
  l[[5]] = anti_join(n.non.eth.summaries, n.eth.summaries, by = "business_id")

  return(l)
}
```

Let's look at the distribution of categories
```{r category utility}
MakeCategoriesDict = function(categories) {
  category.map = hash()
  # category.map[["key"]]
  for (category.string in categories) {
    str = gsub("\'|\\[|\\]", '', category.string)
    category.list = unlist(strsplit(str, ", "))
    for (category in category.list) {
      count = category.map[[category]]
      if (is.null(count)) {
        category.map[[category]] = 1
      } else {
        category.map[[category]] = count + 1
      }
    }
  }
  del('Restaurants', category.map)  # All the Restaurants have this category, so it's useless.
  return(category.map)
}
```

What you want a bar chart for latinos and non latinos
category is x axis 

Top 50 restaurants among Latinos, based on number of reviews
```{r}
n = 50
latinos = CompileStats(latino.names, n)
latinos[[1]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[1]]$categories)), decreasing = TRUE)
```

Top 50 restaurants among non-Latinos, based on number of reviews
```{r}
latinos[[2]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[2]]$categories)), decreasing = TRUE)
```

Comparing the two sets, which restaurants are in the Top 50 for both groups?
```{r}
latinos[[3]] %>% select(name, num.reviews)
```
Surprisingly, these two groups share 39/50 (78%) of the same restaurants.

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[3]]$categories)), decreasing = TRUE)
```

Restaurants that are popular among Latinos but not among non-Latinos:
```{r}
latinos[[4]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[4]]$categories)), decreasing = TRUE)
```
Not surprisingly, Mexican restaurants show up the most here.

Vice-versa:
```{r}
latinos[[5]] %>% select(name, num.reviews)
```

```{r}
sort(values(MakeCategoriesDict(latinos[[5]]$categories)), decreasing = TRUE)
```
No Mexican restaurants here.

## Similar analysis but for East Asians

Top 50 restaurants among East Asians, based on number of reviews:
```{r}
asians = CompileStats(east.asian.names, n)
asians[[1]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[1]]$categories)), decreasing = TRUE)
```

Top 50 restaurants among non-East Asians:
```{r}
asians[[2]] %>% select(name, num.reviews)
```
20 of these are Asian restaurants.

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[2]]$categories)), decreasing = TRUE)
```
Whereas 11 of these are Asian.

Comparing the two sets, which restaurants are in the Top 50 for both groups?
```{r}
asians[[3]] %>% select(name, num.reviews)
```
33/50 are shared between the sets. 

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[3]]$categories)), decreasing = TRUE)
```

Restaurants that are popular among Asians but not among non-Asians:
```{r}
asians[[4]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[4]]$categories)), decreasing = TRUE)
```

Vice-versa:
```{r}
asians[[5]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[5]]$categories)), decreasing = TRUE)
```
