---
title: "yelp_reviews"
author: "Huey"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(stringi)
library(hash)

setwd("~/code/yelp")
```

# Introduction 
Are there differences in rating restaurants among different ethnicities? For 
instance, perhaps Chinese people rate Chinese restaurants differently from other
Americans (presumably based on authenticity). If so, building a "racist Yelp" could be useful 
for identifying more authetnic restaurants.

Here we look at the Yelp dataset to try to tease out any differences, looking
at Mexican and East Asian restaurants.

### Data
The Yelp data contains 2.2M reviews and 591K tips by 552K users for 77K businesses.

The cities included are Edinburgh (UK), Karlsruhe (Germany), Montreal and Waterloo (Canada),
and Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison (US).

I used the JSON to CSV converter in https://github.com/Yelp/dataset-examples 
to make it easier to load.

We use a list of popular Latino names pulled from http://www.1happybirthday.com/popular_names_latino.php.
We classify a user as Latino if their name appears in this list.

```{r load data}
businesses = read_csv("data/yelp_academic_dataset_business.csv")
reviews = read_csv("data/yelp_academic_dataset_review.csv")
users = read_csv("data/yelp_academic_dataset_user.csv")
users$name = tolower(users$name)

latino.names = read.csv("popular_names_latino.csv", 
                        encoding = "latin1", stringsAsFactors = FALSE)
latino.names$Male.Name = stri_trans_general(latino.names$Male.Name, 
                                            'latin-ascii')
latino.names$Female.Name = stri_trans_general(latino.names$Female.Name, 
                                              'latin-ascii')

latino.names = as.data.frame(c(latino.names$Male.Name, latino.names$Female.Name), stringsAsFactors = FALSE)
names(latino.names) = "name"

latino.names.exclude = read.csv("popular_names_latino_exclude.csv", stringsAsFactors = FALSE)

latino.names = anti_join(latino.names, latino.names.exclude, by = "name")
latino.names$name = tolower(latino.names$name)
```

When classifying ethnicity, it's good to get a rough sense of how many users were
classified as Latino or not:

```{r latino users}
latino.users = semi_join(users, latino.names, by = "name")
non.latino.users = anti_join(users, latino.names, by = "name")

print(paste("Latino:", nrow(latino.users)))
print(paste("Non-Latino:" , nrow(non.latino.users)))
```

Less than 3% of users in the dataset have been classified as Latino. There are 
probably more Latino users in the dataset who have names of European origin, but
since Yelp only gives us access to first names, this is probably the best we can
do at the moment.

Below is a function to get all the top n restaurants of a certain ethnicity and
bucket the user reviews into members of that ethnicity or not. We then run a 
t-test applying the Bonferroni correction to find which differences are 
statistically significant. 

```{r utility function}
GetRestaurantsWithDiffRatings = function(ethnicity, n.restaurants, eth.names) {
  eth.restaurants = businesses %>% filter(grepl(ethnicity, categories)) %>% filter(grepl('Restaurants', categories)) %>% arrange(desc(review_count))
  restaurants = eth.restaurants[1:n.restaurants, ]
  restaurant.reviews = inner_join(restaurants, reviews, by = "business_id")
  restaurant.reviews.users = inner_join(restaurant.reviews, users, by = "user_id")
  
  p.value.cutoff = 0.05 / n.restaurants

  restaurant_names = c()
  eth.ratings = c()
  non.eth.ratings = c()
  num.eth.reviews = c()
  num.non.eth.reviews = c()

  for (id in restaurants$business_id) {
    subset = filter(restaurant.reviews.users, business_id == id)
    eth.reviews = semi_join(subset, eth.names, c("name.y" = "name"))
    non.eth.reviews = anti_join(subset, eth.names, c("name.y" = "name"))
    
    if (length(eth.reviews$stars.y) <= 1 || length(non.eth.reviews$stars.y) <= 1) next  # Prevent errors to t-test
    
    try({
      alt = "two.sided"
      eth.mean = mean(eth.reviews$stars.y)
      non.eth.mean = mean(non.eth.reviews$stars.y)
      if (eth.mean < non.eth.mean) {
        alt = "less"
      }
      if (eth.mean > non.eth.mean) {
        alt = "greater"
      }
      res = t.test(eth.reviews$stars.y, non.eth.reviews$stars.y, alternative = alt)
      if (res$p.value < p.value.cutoff) {
        restaurant_names = c(restaurant_names, subset$name.x[1])
        eth.ratings = c(eth.ratings, eth.mean)
        non.eth.ratings = c(non.eth.ratings, non.eth.mean)
        num.eth.reviews = c(num.eth.reviews, nrow(eth.reviews))
        num.non.eth.reviews = c(num.non.eth.reviews, nrow(non.eth.reviews))
      }
    }
    )
  }
  return(data.frame(restaurant_names, eth.ratings, non.eth.ratings, num.eth.reviews, num.non.eth.reviews))
}
```

## Mexican Restaurants
The following table shows the restaurants where Latino ratings and non-Latino ratings
had statistically significant differences:

```{r latino data frame}
results = GetRestaurantsWithDiffRatings('Mexican', 100, latino.names)
results
```

For all of these, the sample size of Latino reviews is really small. It seems that
building a "racist Yelp" for Mexican restaurants would not be all that useful.

## Chinese Restaurants
To get Chinese-sounding names, we will compile a list of all pinyin sounds and
see if they exist in any first names. This runs in quadratic time, so we save
the users who matched.

```{r chinese users}

load(file = "chinese.users.RData")

if (is.null(chinese.users)) {
  pinyin = read.csv("pinyin_sounds.csv", stringsAsFactors = FALSE)
  pinyin.pattern = paste(pinyin$sound, collapse ="|")
  
  exclude = "benji|ique|cque|arjun|axi|lexi|xine|jinx|quant|change|strong|cruise|chunky|suite|burgh fan|tx fan|eping|pping|lping"
  
  tictoc::tic()
  indexes = vector("logical", length = nrow(users))
  for (i in 1:nrow(users)) {
    if (length(grep(exclude, users$name[i])) > 0) next
    
    matches = grep(pinyin.pattern, users$name[i])
    indexes[i] = length(matches) > 0
  }
  tictoc::toc()
  chinese.users = users[indexes,]
  nrow(chinese.users)
  
  # save chinese.users in case it takes a long time
  save(chinese.users, file = "chinese.users.RData")
}

chinese.names = as.data.frame(unique(chinese.users$name), stringsAsFactors = FALSE)
names(chinese.names) = "name"
set.seed(1)
sample_n(chinese.names, 10)
```

There are 7755 users classified as Chinese, or about 1.4% of the user population.
Sanity checking: Americans of Chinese descent consititute 1.2% of the total U.S.
popuplation as of 2010. 

You can see from the sample that some of the names look more obviously Chinese than not.
This suggests that we probably classified too many names as Chinese,
as it is likely that many Chinese users of Yelp have Western first names.

```{r}
results = GetRestaurantsWithDiffRatings('Chinese', 100, chinese.names)
names(results) = c("name", "chinese_ratings", "non_chinese_ratings", "n_chinese", "n_non_chinese")
results
```

Only 3 restaurants have statistically different ratings, and two of them only have
2 Chinese reviews!

## Chinese Names Take 2
An alternative is to use the baby names from the following site:
http://www.top-100-baby-names-search.com/chinese-name-meanings.html
```{r chinese users take 2}
chinese.names2 = read.csv("top_100_baby_names_chinese.csv", stringsAsFactors = FALSE)
chinese.names2$name = tolower(chinese.names2$name)
chinese.names2 = chinese.names2[-124,]  # remove "juan"
chinese.users2 = semi_join(users, chinese.names2, by = "name")

results = GetRestaurantsWithDiffRatings('Chinese', 100, chinese.names2)
names(results) = c("name", "chinese_ratings", "non_chinese_ratings", "n_chinese", "n_non_chinese")
results
```

This yielded 2028 users and a different set of restaurants that had statistically
significantly different ratings. However, the number of reviews by Chinese people
are still too low to make any significant conclusions.

## East Asian Restaurants
What if we expand our search to Korean + Japanese + Chinese? Here we just use the 
baby names list for Chinese names.

```{r}
korean.names = read.csv("top_100_baby_names_korean.csv", stringsAsFactors = FALSE)
korean.names$name = tolower(korean.names$name)

japanese.names = read.csv("top_100_baby_names_japanese.csv", stringsAsFactors = FALSE)
japanese.names$name = tolower(japanese.names$name)

korean.users = semi_join(users, korean.names, by = "name")
japanese.users = semi_join(users, japanese.names, by = "name")

east.asian.names = as.data.frame(c(chinese.names2$name, japanese.names$name, korean.names$name), stringsAsFactors = FALSE)
names(east.asian.names) = c("name")
east.asian.users = semi_join(users, east.asian.names, by = "name")

results = GetRestaurantsWithDiffRatings('Chinese|Korean|Japanese', 100, east.asian.names)
names(results) = c("name", "asian_ratings", "non_asian_ratings", "n_asian", "n_non_asian")
results
```

The only restaurant that has more than 10 East Asian reviews is Gangnam Asian
BBQ Dining. Its rating among Asians is slightly higher.

## East Asian Users using pinyin for Chinese names
```{r east asian restaurants}
east.asian.names = as.data.frame(c(chinese.names$name, japanese.names$name, korean.names$name), stringsAsFactors = FALSE)
names(east.asian.names) = c("name")
east.asian.users = semi_join(users, east.asian.names, by = "name")

results = GetRestaurantsWithDiffRatings('Chinese|Korean|Japanese', 100, east.asian.names)
results
```

Again, nothing terribly interesting here.

# Conclusion
You can't construct a good "racist Yelp" from the data provided. If we had a better
way of classifying the race of users, we might be able to get better results.

## Restaurant preferences
What kind of restaurants do Latinos review?
What kind of restaurants do Asians review?

What's the distribution of review count on yelp?
```{r distribution of review count}
bw = function(x) diff(range(x)) / (2 * IQR(x) / length(x)^(1/3))
restaurants = businesses %>% filter(grepl('Restaurants', categories)) %>% arrange(desc(review_count))
ggplot(restaurants, aes(x = review_count)) + geom_histogram()
```

This seems in line with what we expect. The vast majority of restaurants have 
a few reviews, with a long tail -- some restaurants have many more reviews.

Here is some code to group reviews by ethnicity.
```{r review counts for ethnicity}
ReviewCountsForEthnicity = function(eth.names) {
  eth.users = semi_join(users, eth.names, by = "name")
  non.eth.users = anti_join(users, eth.names, by = "name")
  
  restaurant.reviews = inner_join(restaurants, reviews, by = "business_id")
  restaurant.reviews.eth.users = inner_join(restaurant.reviews, eth.users, by = "user_id")
  restaurant.reviews.non.eth.users = inner_join(restaurant.reviews, non.eth.users, by = "user_id")
  
  by.restaurant = group_by(restaurant.reviews.eth.users, business_id)
  eth.summary = summarise(by.restaurant, num.reviews = n()) 
  eth.summary = inner_join(eth.summary, restaurants, by = "business_id") %>% select(business_id, name, num.reviews, categories) 
  
  by.restaurant = group_by(restaurant.reviews.non.eth.users, business_id)
  non.eth.summary = summarise(by.restaurant, num.reviews = n()) %>% arrange(desc(num.reviews))
  non.eth.summary = inner_join(non.eth.summary, restaurants, by = "business_id") %>% select(business_id, name, num.reviews, categories) 
  
  return(list(eth.summary, non.eth.summary))
}
```

```{r compile stats}
CompileStats = function(names, n) {
  l = list()
  summaries = ReviewCountsForEthnicity(names)
  
  n.eth.summaries = head(summaries[[1]] %>% arrange(desc(num.reviews)), n)
  l[[1]] = n.eth.summaries
  
  n.non.eth.summaries = head(summaries[[2]] %>% arrange(desc(num.reviews)) , n)
  l[[2]] = n.non.eth.summaries
  
  l[[3]] = semi_join(n.eth.summaries, n.non.eth.summaries, by = "business_id")
  l[[4]] = anti_join(n.eth.summaries, n.non.eth.summaries, by = "business_id")
  l[[5]] = anti_join(n.non.eth.summaries, n.eth.summaries, by = "business_id")

  return(l)
}
```

This is a utlity function to count the categories of a set of restaurants.
```{r category utility}
MakeCategoriesDict = function(categories) {
  category.map = hash()
  # category.map[["key"]]
  for (category.string in categories) {
    str = gsub("\'|\\[|\\]", '', category.string)
    category.list = unlist(strsplit(str, ", "))
    for (category in category.list) {
      count = category.map[[category]]
      if (is.null(count)) {
        category.map[[category]] = 1
      } else {
        category.map[[category]] = count + 1
      }
    }
  }
  del('Restaurants', category.map)  # All the Restaurants have this category, so it's useless.
  return(category.map)
}
```

Top 100 restaurants among Latinos, based on number of reviews:
```{r}
n = 100
latinos = CompileStats(latino.names, n)
latinos[[1]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[1]]$categories)), decreasing = TRUE)
```
Mostly Nightlife, Bars, American Food.

Top 100 restaurants among non-Latinos, based on number of reviews
```{r}
latinos[[2]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[2]]$categories)), decreasing = TRUE)
```
Mostly Nightlife, Bars, American Food.

Comparing the two sets, which restaurants are in common?
```{r}
latinos[[3]] %>% select(name, num.reviews)
```
75%!

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[3]]$categories)), decreasing = TRUE)
```
Bars, Nightlife, American food.

Restaurants that are popular among Latinos but not among non-Latinos:
```{r}
latinos[[4]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(latinos[[4]]$categories)), decreasing = TRUE)
```
Hard to see, a little bit of Mexican food.

Vice-versa:
```{r}
latinos[[5]] %>% select(name, num.reviews)
```

```{r}
sort(values(MakeCategoriesDict(latinos[[5]]$categories)), decreasing = TRUE)
```
No Mexican restaurants here.

## Similar analysis but for East Asians

Top 100 restaurants among East Asians, based on number of reviews:
```{r}
asians = CompileStats(east.asian.names, n)
asians[[1]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[1]]$categories)), decreasing = TRUE)
```
Japanese, Buffets, Bars, Chinese, Nightlife

Top 100 restaurants among non-East Asians:
```{r}
asians[[2]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[2]]$categories)), decreasing = TRUE)
```
Bars, Nightlife, American

Comparing the two sets, which restaurants are in the Top 50 for both groups?
```{r}
asians[[3]] %>% select(name, num.reviews)
```
55% are shared amongst the two groups. This is different than what we saw with
the Latino vs. non-Latino comparisons above.

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[3]]$categories)), decreasing = TRUE)
```
Bars, Buffets, Nightlife

Restaurants that are popular among Asians but not among non-Asians:
```{r}
asians[[4]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[4]]$categories)), decreasing = TRUE)
```
Asian food

Vice-versa:
```{r}
asians[[5]] %>% select(name, num.reviews)
```

Category counts:
```{r}
sort(values(MakeCategoriesDict(asians[[5]]$categories)), decreasing = TRUE)
```
Non-Asian food. It is interesting that there is more disparity between Asians and non-Asians than Latinos and non-Latinos.