---
title: "yelp_reviews"
author: "Huey"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(stringi)

setwd("~/code/yelp")
```

Hypotheses:

* Latinos are more critical of Mexican restaurants than non-Latinos.
* Latinos will like more authentic Mexican restaurants than non-Latinos.

## Load businesses
First, let's load the Yelp businesses. I used the JSON to CSV converter in https://github.com/Yelp/dataset-examples to make it easier for me to load.

```{r load data}
businesses = read_csv("data/yelp_academic_dataset_business.csv")
reviews = read_csv("data/yelp_academic_dataset_review.csv")
users = read_csv("data/yelp_academic_dataset_user.csv")

latino_names = read.csv("popular_names_latino.csv", 
                        encoding = "latin1", stringsAsFactors = FALSE)
latino_names$Male.Name = stri_trans_general(latino_names$Male.Name, 
                                            'latin-ascii')
latino_names$Female.Name = stri_trans_general(latino_names$Female.Name, 
                                              'latin-ascii')

latino_names = as.data.frame(c(latino_names$Male.Name, latino_names$Female.Name), stringsAsFactors = FALSE)
names(latino_names) = "name"
```

## Get Mexican restaurants
Get Mexican restaurants with the most reviews. 
```{r}
mexican = businesses %>% filter(grepl('Mexican', categories)) %>% filter(grepl('Restaurants', categories)) %>% arrange(desc(review_count))

restaurants = mexican[1,]
restaurants$review_count
```

## Looking at the top restaurant, Tacos El Gordo
Get the reviews for these restaurants
```{r join reviews and restaurants}
restaurant_reviews = inner_join(restaurants, reviews, by = "business_id")
dim(restaurant_reviews)
```

Get the user names associated with these reviews
```{r}
restaurant_reviews_users = inner_join(restaurant_reviews, users, by = "user_id")
dim(restaurant_reviews_users)  
```

Bucket into Latino and non-Latino
```{r}
latino_reviews = semi_join(restaurant_reviews_users, latino_names, c("name.y" = "name"))
non_latino_reviews = anti_join(restaurant_reviews_users, latino_names, c("name.y" = "name"))
```

Compare distribution of review scores
```{r}
summary(latino_reviews$stars.y)
summary(non_latino_reviews$stars.y)
```
It seems like the mean of latino reviews is lower, 4.212 to 4.322.

Plot them:
```{r}
stars = c(latino_reviews$stars.y, non_latino_reviews$stars.y)
latin = c(rep("Latin", length(latino_reviews$stars.y)), rep("Non-Latin", length(non_latino_reviews$stars.y)))
reviews_latin_or_non = data.frame(stars, latin)

ggplot(reviews_latin_or_non, aes(x = stars)) + geom_bar() + facet_wrap(~latin)
```

It's actually hard to tell if there's a difference, since there are way more
reviews classified as non-Latin.

t-test
```{r}
t.test(latino_reviews$stars.y, non_latino_reviews$stars.y)
t.test(latino_reviews$stars.y, non_latino_reviews$stars.y, alternative = "less")
```
Not statistically significant.

## List of Questions
Is there a Chipotle in this dataset? It would be interesting to see if there
are differences.